Learning a functional distance between organizations with unsupervised embeddings of scientific mobility

### 2020-07-20 paper/current/main.tex
remove "elite" naming

The geographic structure of the global embedding is mirrored in the re-projection of the United States, wherein organizations are largely organized according to regions and states (Fig.~\ref{fig:projection}d).
By re-projecting Massachusetts (Fig.~\ref{fig:projection}e), we observe structure based on urban centers (Boston vs. Worcester) as well as on organization type (e.g., hospitals vs. universities) and university systems.
For example, even though UMass Boston is located in Boston, it clusters with  other universities in the UMass System ($d = 0.34$) rather than the other typically more elite and research-focused universities in Boston ($d = 0.41$), reflecting a relative lack of mobility between the two systems, even within the same state.
Similar structures can be observed in other states such as among New York's CUNY and SUNY systems (Fig.~\ref{fig:supp:proj_newyork}), Pennsylvania's state system (Fig.~\ref{fig:supp:proj_pennsylvania}), Texas's Agricultural and Mechanical universities (Fig.~\ref{fig:supp:proj_texas}), and between the University of California and State University of California systems (Fig.~\ref{fig:supp:proj_california}).

@@

The embedding space captures prestige hierarchies that can be used to reconstruct academic rankings. 
We derive a ranking using the SemAxis prestige axis in Fig.~\ref{fig:semaxis}a and compare them to traditional university rankings.
The SemAxis ranking is defined from the position of organizations along the prestige axis, with the top-ranked being the furthest towards the \textit{Elite} end of the axis.
We find that the SemAxis ranking is correlated with the Times ranking (Spearman's $\rho = 0.73$, Fig.~\ref{fig:semaxis}b).
The strongest outliers that were ranked more highly in the Times Ranking than the SemAxis ranking tend to be large state universities such as Arizona State University and  University of Florida;
Those ranked higher in the SemAxis ranking tend to be relatively-smaller universities such as the University of San Francisco and University of Maryland Baltimore County that are close to other prestigious universities.

@@

Scientific mobility is complex and can be defined in many possible ways~\autocite{robinson2019mobility} and at multiple levels of analysis.
Here, authors are classified as mobile when they have at least two distinct organization ids in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.
Under this definition, mobile authors constitute 3,007,192 or 23.2 percent of all authors and 17,700,095 author-affiliation combinations.
Mobile authors were associated with, on average, 2.5 distinct organizational affiliations.
Rates of mobility tended to differ across countries.
For example, France, Qatar, the USA, Iraq, and Luxembourg had the most mobile authors (Fig.~\ref{fig:supp:descriptives}c).
However, due to their size, the USA, accounted for nearly 40 percent of all mobile authors worldwide (Fig.~\ref{fig:supp:descriptives}a), with 10 countries accounting for 80 percent of all mobility (Fig.~\ref{fig:supp:descriptives}b).
The countries with the highest proportion of mobile scientists are France, Qatar, the United States, and Iraq, whereas those with the lowest are Jamaica, Serbia, Bosnia \& Herzegovina, and North Macedonia\ref{fig:supp:descriptives}.c).
In most cases, countries with a high degree of inter-organization mobility also have a high degree of international mobility, indicating a high proportion of their total mobility being international (Fig.~\ref{fig:supp:descriptives}.d);
However, some countries such as France and the United States seem to have more domestic mobility than international mobility. 
While the number of publications has increased year-to-year, the mobility and disciplinary makeup of the dataset has not significantly changed across the period of study (Fig.~\ref{fig:supp:pubs_over_time}).

>>>>

The geographic structure of the global embedding is mirrored in the re-projection of the United States, wherein organizations are largely organized according to regions and states (Fig.~\ref{fig:projection}d).
By re-projecting Massachusetts (Fig.~\ref{fig:projection}e), we observe structure based on urban centers (Boston vs. Worcester) as well as on organization type (e.g., hospitals vs. universities) and university systems.
For example, even though UMass Boston is located in Boston, it clusters with  other universities in the UMass System ($d = 0.34$) rather than the other typically more highly-ranked and research-focused universities in Boston ($d = 0.41$), reflecting a relative lack of mobility between the two systems, even within the same state.
Similar structures can be observed in other states such as among New York's CUNY and SUNY systems (Fig.~\ref{fig:supp:proj_newyork}), Pennsylvania's state system (Fig.~\ref{fig:supp:proj_pennsylvania}), Texas's Agricultural and Mechanical universities (Fig.~\ref{fig:supp:proj_texas}), and between the University of California and State University of California systems (Fig.~\ref{fig:supp:proj_california}).

@@

The embedding space captures prestige hierarchies that can be used to reconstruct academic rankings. 
We derive a ranking using the SemAxis prestige axis in Fig.~\ref{fig:semaxis}a and compare them to traditional university rankings.
The SemAxis ranking is defined from the position of organizations along the prestige axis, with the top-ranked being the furthest towards the high-ranked end of the axis.
We find that the SemAxis ranking is correlated with the Times ranking (Spearman's $\rho = 0.73$, Fig.~\ref{fig:semaxis}b).
The strongest outliers that were ranked more highly in the Times Ranking than the SemAxis ranking tend to be large state universities such as Arizona State University and  University of Florida;
Those ranked higher in the SemAxis ranking tend to be relatively-smaller universities such as the University of San Francisco and University of Maryland Baltimore County that are close to other prestigious universities.

@@

Scientific mobility is complex and can be defined in many possible ways~\autocite{robinson2019mobility} and at multiple levels of analysis.
Here, authors are classified as mobile when they have at least two distinct organization IDs in their trajectory, meaning that they have published using two or more distinct affiliations between 2008 and 2019.
Under this definition, mobile authors constitute 3,007,192 or 23.2 percent of all authors and 17,700,095 author-affiliation combinations.
Mobile authors were associated with, on average, 2.5 distinct organizational affiliations.
Rates of mobility tended to differ across countries.
For example, France, Qatar, the USA, Iraq, and Luxembourg had the most mobile authors (Fig.~\ref{fig:supp:descriptives}c).
However, due to their size, the USA, accounted for nearly 40 percent of all mobile authors worldwide (Fig.~\ref{fig:supp:descriptives}a), with 10 countries accounting for 80 percent of all mobility (Fig.~\ref{fig:supp:descriptives}b).
The countries with the highest proportion of mobile scientists are France, Qatar, the United States, and Iraq, whereas those with the lowest are Jamaica, Serbia, Bosnia \& Herzegovina, and North Macedonia\ref{fig:supp:descriptives}.c).
In most cases, countries with a high degree of inter-organization mobility also have a high degree of international mobility, indicating a high proportion of their total mobility being international (Fig.~\ref{fig:supp:descriptives}.d);
However, some countries such as France and the United States seem to have more domestic mobility than international mobility. 
While the number of publications has increased year-to-year, the mobility and disciplinary makeup of the dataset has not significantly changed across the period of study (Fig.~\ref{fig:supp:pubs_over_time}).
